### 2015-06-24 Dissertation/chapter-introduction.tex

\chapter{Introduction}
\label{chap-introduction}
A primary purpose of evolutionary optimization is to efficiently find good solutions
to challenging real world problems with minimal prior knowledge about the problem itself.
This driving goal has created search algorithms that can escape user bias to create
truly novel results, sometimes publishable or patentable in their own right~\cite{kannappan:2014:humies}.
While it is not possible for any algorithm to do better than random search across all possible
problems~\cite{Wolpert:1997:nfl}, effectiveness can be achieved by assuming the search
landscape has structure and then biasing the algorithm toward exploiting that structure.

>>>>

\chapter{Introduction}
\label{chap-introduction}
A primary purpose of evolutionary optimization is to efficiently find good solutions
to challenging real-world problems with minimal prior knowledge about the problem itself.
This driving goal has created search algorithms that can escape user bias to create
truly novel results, sometimes publishable or patentable in their own right~\cite{kannappan:2014:humies}.
While it is not possible for any algorithm to do better than random search across all possible
problems~\cite{Wolpert:1997:nfl}, effectiveness can be achieved by assuming the search
landscape has structure and then biasing the algorithm toward exploiting that structure.

### 2015-06-24 Dissertation/chapter-proposed.tex
Removed "shall", tried to fix effect/affect

\chapter{Proposed Research}
The Parameter-less Population Pyramid (P3) was shown to be effective in Chapter~\ref{chap-benchmarking}
when applied to black-box optimization. Some of this effectiveness in due to P3's integration of
efficient local search and its ability to automatically detect and exploit variable epistasis.
In Chapter~\ref{chap-graybox} we presented the gray-box domain, which contains many interesting
real world problems, and the improvements that can be made to local search in this domain.
This presents the following questions:

@@

We shall tentatively refer to the first proposed algorithm of this type as the Simple Uniform crossover Tournament (SUT)
algorithm. Conceptually, SUT is meant to be the simplest way to combine uniform crossover with hamming ball hill climbing.
Solutions compete in an elimination tournament. Each contest in the tournament uses crossover between the solutions, with each
offspring subsequently brought to a local optimum using hamming ball hill climbing. The best solution of the parents and offspring then
advances up the tournament.

@@

As an initial experiment, we have focused on recreating the P3 style of linkage tree without using entropy tables.
Figure~\ref{fig-sfx-tree} provides an algorithm for using subfunction information to create a linkage tree.
Initially all variables start in their own cluster. Each subfunction is then used in a random order such that
if that subfunction overlaps multiple top level clusters it merges them to create a new cluster.

@@

\subsection{Preliminary Empirical Comparison}
We tested P3 using the gray-box \Call{SubfunctionTree}{} in place of the black-box \Call{Cluster-Creation}{}
and using the gray-box hamming ball hill climber instead of black-box hill climbing. As an added efficiency
instead of recalculating all $delta$ values and assuming all moves are open each time a crossover donation is made,
only $delta$s and moves effected by the donation are updated. This means that hill climbing requires $O(|cluster|+I)$
time.

@@

Beyond the randomly generated instances, we plan to seek out real world problems that can be expressed in this domain.
This will help solidify the conclusions about the applicability of these methods outside of the optimization community.
Similarly, we plan to compare the results of our gray-box methods with algorithms specifically designed for each problem
domain. This provides an interesting contention between the time required to develop problem specific solvers and the
quality of using gray-box methods.

As the hamming ball hill climber has a user specified parameter $r$, we will need to explore this parameter space.
Specifically we will want to ensure all comparisons use an appropriate value for $r$ that does not scale with
$N$. This process will likely involve brute force testing all small $r$ values for moderately large problems in each class
and then choosing the one that performs best. That $r$ value will then be used for all problem sizes for that problem class.

@@

\section{Summary of Proposed Research Questions}
\begin{itemize}
\item How does the combination of uniform crossover with hamming ball hill climbing compare to hill climbing alone?
\item Does preserving subfunctions intact during crossover improve search efficiency?
\item How does black-box P3 compare with gray-box P3?
\item How do the more generic gray-box methods compare with problem specific methods for solving real world problems?
\end{itemize}

>>>>

\chapter{Proposed Research}
The Parameter-less Population Pyramid (P3) was shown to be effective in Chapter~\ref{chap-benchmarking}
when applied to black-box optimization. Some of this effectiveness in due to P3's integration of
efficient local search and its ability to automatically detect and exploit variable epistasis.
In Chapter~\ref{chap-graybox} we presented the gray-box domain, which contains many interesting
real-world problems, and the improvements that can be made to local search in this domain.
This presents the following questions:

@@

We tentatively refer to the first proposed algorithm of this type as the Simple Uniform crossover Tournament (SUT)
algorithm. Conceptually, SUT is meant to be the simplest way to combine uniform crossover with hamming ball hill climbing.
Solutions compete in an elimination tournament. Each contest in the tournament uses crossover between the solutions, with each
offspring subsequently brought to a local optimum using hamming ball hill climbing. The best solution of the parents and offspring then
advances up the tournament.

@@

As an initial experiment, we have focused on recreating the P3 style of linkage tree without using entropy tables.
Figure~\ref{fig-sfx-tree} provides an algorithm for using subfunction information to create a linkage tree.
Initially all variables start in their own cluster. Each subfunction is then used in a random order such that
if that subfunction overlaps multiple top-level clusters it merges them to create a new cluster.

@@

\subsection{Preliminary Empirical Comparison}
We tested P3 using the gray-box \Call{SubfunctionTree}{} in place of the black-box \Call{Cluster-Creation}{}
and using the gray-box hamming ball hill climber instead of black-box hill climbing. As an added efficiency
instead of recalculating all $delta$ values and assuming all moves are open each time a crossover donation is made,
only $delta$s and moves affected by the donation are updated. This means that hill climbing requires $O(|cluster|+I)$
time.

@@

Beyond the randomly generated instances, we plan to seek out real-world problems that can be expressed in this domain.
This will help solidify the conclusions about the applicability of these methods outside of the optimization community.
Similarly, we plan to compare the results of our gray-box methods with algorithms specifically designed for each problem
domain. This provides an interesting contention between the time required to develop problem-specific solvers and the
quality of using gray-box methods.

As the hamming ball hill climber has a user-specified parameter $r$, we will need to explore this parameter space.
Specifically we will want to ensure all comparisons use an appropriate value for $r$ that does not scale with
$N$. This process will likely involve brute force testing all small $r$ values for moderately large problems in each class
and then choosing the one that performs best. That $r$ value will then be used for all problem sizes for that problem class.

@@

\section{Summary of Proposed Research Questions}
\begin{itemize}
\item How does the combination of uniform crossover with hamming ball hill climbing compare to hill climbing alone?
\item Does preserving subfunctions intact during crossover improve search efficiency?
\item How does black-box P3 compare with gray-box P3?
\item How do the more generic gray-box methods compare with problem-specific methods for solving real-world problems?
\end{itemize}
